### 2022-11-30 code/3_analysis/3_analysis_direction.Rmd
added a lot of analysis. prepared to move to coglink_cogsci

```{r combine corpus and behavior data, include = F}
source("../2_combine/2_combine.R")
df.basefreq <- read.csv("../../data/corpus/basefreq_EMV.csv")

@@

df <- left_join(df, df.basefreq, by = "triad")
```

## Does cue given match  predict behavior better than match given cue?

## Do we extend previous work reporting a preference for taxonomic matching in the US and thematic matching in Asia?

```{r warning=FALSE, include=FALSE, paged.print=TRUE}
df.country <- df %>%
  group_by(subject, country) %>%
  summarize(theme_resp_percent = mean(responses_theme, na.rm = T))

df.country_sum <- df.country %>%
  group_by(country) %>%
  summarize(mean_theme_resp_percent = mean(theme_resp_percent), 
            sd_theme_resp_percent = sd(theme_resp_percent))

fit.country = glmer(responses_theme ~ country + (1 | subject) + (country | triad), 
                    data = df, 
                    family = "binomial")
summary(fit.country) 
fit.country.anova = Anova(fit.country, type=3)

fit.country_EN_VN = glmer(responses_theme ~ country + (1 | subject) + (country | triad), 
                    data = df %>% filter(country != "China"),
                    family = "binomial")
summary(fit.country_EN_VN) 
```
The group means of proportion of thematic response in mainland China is the highest (M = `r round((df.country_sum %>% filter(country == "China"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "China"))$sd_theme_resp_percent, 2)`), followed by the groups means in Vietnam (M = `r round((df.country_sum %>% filter(country == "Vietnam"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "Vietnam"))$sd_theme_resp_percent, 2)`), which is slightly higher than that of the US (M = `r round((df.country_sum %>% filter(country == "US"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "US"))$sd_theme_resp_percent, 2)`).

```{r echo=FALSE, warning=FALSE, fig.cap="Proportion of thematic responses by country. Only US-China responding comparison shows a siginficant difference. We could not extend this to US-Vietnam responding comparison."}

#show violin plot
ggplot(df.country,
       mapping = aes(x = factor(country, levels=c("China", "US", "Vietnam")), 
                     y = theme_resp_percent, 
                     color = factor(country, levels=c("China", "US", "Vietnam")))) +
  geom_violin() +
  geom_jitter(height = 0, 
              alpha = 0.3) +  
  stat_summary(fun.data = "mean_cl_boot", 
               geom = "pointrange") + 
  labs(y = "Proportion Thematic Chosen", 
       x = "Country", 
       color = "Country") + 
  scale_color_manual(values=c("#D63230", "#1C77C3", "#F39237"))
```

```{r correlation of country response, echo=FALSE, warning=FALSE}
#US-China, US-Vietnam, China-Vietnam
df.triad_country <- df %>%
  group_by(triad, country) %>%
  summarize(emp_theme_prop = mean(responses_theme))

df.triad_country <- pivot_wider(data = df.triad_country, 
                                names_from = country, 
                                values_from = emp_theme_prop)
plotUSCN = ggplot(data = df.triad_country,
                  mapping = aes(x = US, 
                                y = China, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)", 
        y = "Proportion Thematic Chosen (China)") + 
   xlim(0, 1) + 
   ylim(0, 1) + 
  theme(axis.title = element_text(size=rel(0.75)))

plotUSVN = ggplot(data = df.triad_country,
                  mapping = aes(x = US, 
                                y = Vietnam, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)", 
        y = "Proportion Thematic Chosen (Vietnam)") + 
   xlim(0, 1) + 
   ylim(0, 1)  + 
  theme(axis.title = element_text(size=rel(0.75)))

plotCNVN = ggplot(data = df.triad_country,
                  mapping = aes(x = Vietnam,
                                y = China, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (Vietnam)", 
        y = "Proportion Thematic Chosen (China)") + 
   xlim(0, 1) + 
   ylim(0, 1)  + 
  theme(axis.title = element_text(size=rel(0.75)))

corr.USCN = apa_print(cor.test(df.triad_country$US, 
                  df.triad_country$China, na.rm=T))

corr.USVN = apa_print(cor.test(df.triad_country$US, 
                  df.triad_country$Vietnam, na.rm=T))

corr.CNVN = apa_print(cor.test(df.triad_country$China, 
                  df.triad_country$Vietnam, na.rm=T))
```

```{r warning = F, fig.width = 8, fig.cap="Correlation between proportion thematic chosen by US-China participants, US-Vietnam participants, and China-Vietnam participants. All three comparisons show strong correlation, suggesting a strong signal of universality across cultural contexts regarding similarity judgments"}
plotUSCN + plotUSVN + plotCNVN +
  plot_layout(ncol = 3)
```

With a Pearson correlation test, we observed high correlations between thematic responding across all three countries. The proportion of US thematic responding is strongly correlated with thematic responding in China, `r corr.USCN$full_result`. The proportion of US thematic responding is strongly correlated with thematic responding in Vietnam, `r corr.USVN$full_result`. Likewise, the proportion of China thematic response is strongly correlated with thematic responding in Vietnam, `r corr.CNVN$full_result`. This points to a strong signal of universality across cultural contexts regarding similarity judgments.

```{r USCN effect size comparison, include=FALSE}
USCN_cohen <- cohen.d(df.triad_country$China, df.triad_country$US)
```

For comparison with previous work, we also included an effect size comparison for the difference in US-China thematic response data (though this was not a planned analysis). The effect size for this analysis was d = `r USCN_cohen$estimate`, a small effect size. This is smaller than the effect size from @Ji2004, even when we are unable to make a direct comparison. Calculating the effect size from the ANOVA reported in @Ji2004 between US participants tested in English and mainland China participants tested in English (both groups tested in the US) yields a Cohen's d = 0.81. @Ji2004 reported that mean thematic response for mainland China participants tested in Mandarin Chinese is higher than for those tested in English, so the effect size for a comparison between mainland China participants tested in Mandarin Chinese and US participants tested in English (parallel to our comparison) would likely have yielded an even larger effect size than 0.814. It is possible that we failed to replicate the magnitude of effect from @Ji2004 because the previous study calculated their dependent variable as frequency of thematic responses minus frequency of taxonomic responses, whereas our dependent variable is proportion of thematic response over total number of responses.

To test for cross-context differences in similarity judgments between the countries, we ran a mixed-effects logistic regression predicting triad responding (taxonomic or thematic) with country (US, China, or Vietnam) as a fixed effect. As random effects, we included an intercept per subject and one per triad, as well as by-triad random slopes for country to account for variation in the country effect across triads. In R syntax, the model is: response ~ country + (1 | subject) + (country | triad).

Overall, there is a significant effect of country on proportion of thematic responses ($\chi^2$(`r fit.country.anova["country", "Df"]`) = `r round(fit.country.anova["country", "Chisq"], 2)`, p < .001). However, this effect is driven by the difference between US and China responding ($\beta$ = `r round(fixef(fit.country)["countryUS"], 2)`, p < .001). There is no statistical difference between the Vietnam and China responding ($\beta$ = `r round(fixef(fit.country)["countryVietnam"], 2)`, p = `r round(as.data.frame(summary(fit.country)[["coefficients"]])["countryVietnam", "Pr(>|z|)"], 2)`), and the US and Vietnam responding ($\beta$ = `r round(fixef(fit.country_EN_VN)["countryVietnam"], 2)`, p = `r round(as.data.frame(summary(fit.country_EN_VN)[["coefficients"]])["countryVietnam", "Pr(>|z|)"], 3)`).

On this analysis, we do not find support that the US-China tendencies toward taxonomic and thematic responding (respectively) extend to the US-Vietnam comparison. Accordingly, we cannot speak to overall biases toward thematic responding across Asian cultural contexts broadly, but we do replicate the differences documented by @Ji2004 between the US and China. However, in our corpus model comparison, we do find evidence for different, more fine-grained variation in similarity judgments between the US and Vietnam.

## Can the differences in similarity judgments between English, Mandarin and Vietnamese speakers be explained by variation in lexical statistics?
### Is responding in each cultural context predicted by the corresponding lexical statistics (raw co-occurrences and cosine distance)? 
```{r Pearson correlation between raw co-occurrences of corpora and responding, warning=FALSE, include=FALSE}

#raw co-occurrences
df.triad_country_freq <- df %>%
  filter(!(triad %in% triads_omit)) %>% # omit triads that do not have meaningful prediction (at least 1 non-zero prediction in both languages)
  group_by(triad, country, theme_freq_prop_E, theme_freq_prop_V, theme_freq_prop_M) %>%  # not working 
  summarize(emp_theme_prop = mean(responses_theme))
df.triad_country_freq   #sanity

df.triad_country_freq_E <- df.triad_country_freq %>%
  filter(country == "US")

#plotting English corpus against US responding data
plotE = ggplot(data = df.triad_country_freq_E,
        mapping = aes(x = emp_theme_prop,
                     y = theme_freq_prop_E,
                     label = triad)) +
   geom_text() +
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)",
        y = "Proportion Thematic Predicted (English)")
plotE

df.triad_country_freq_V <- df.triad_country_freq %>%
  filter(country == "Vietnam")

#plotting Vietnamese corpus against VN responding data
plotV = ggplot(data = df.triad_country_freq_V,
        mapping = aes(x = emp_theme_prop,
                     y = theme_freq_prop_V,
                     label = triad)) +
   geom_text() +
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (Vietnam)",
        y = "Proportion Thematic Predicted (Vietnamese)")
plotV

df.triad_country_freq_M <- df.triad_country_freq %>%
  filter(country == "China")

#plotting China corpus against CN responding data
plotM = ggplot(data = df.triad_country_freq_M,
        mapping = aes(x = emp_theme_prop,
                     y = theme_freq_prop_M,
                     label = triad)) +
   geom_text() +
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (China)",
        y = "Proportion Thematic Predicted (Chinese)")
plotM

#Pearson correlation between corpus and responding
corr.E = apa_print(cor.test(df.triad_country_freq_E$emp_theme_prop,
                  df.triad_country_freq_E$theme_freq_prop_E, na.rm=T))
corr.E

corr.V = apa_print(cor.test(df.triad_country_freq_V$emp_theme_prop,
                  df.triad_country_freq_V$theme_freq_prop_V, na.rm=T))
corr.V

corr.M = apa_print(cor.test(df.triad_country_freq_M$emp_theme_prop,
                  df.triad_country_freq_M$theme_freq_prop_M, na.rm=T))
corr.M

#Pearson correlation between EN and VI corpus
corr.EV = apa_print(cor.test(df.triad_country_freq_V$theme_freq_prop_V,
                   df.triad_country_freq_E$theme_freq_prop_E, na.rm=T))
corr.EM = apa_print(cor.test(df.triad_country_freq_M$theme_freq_prop_M,
                   df.triad_country_freq_E$theme_freq_prop_E, na.rm=T))
corr.MV = apa_print(cor.test(df.triad_country_freq_V$theme_freq_prop_V,
                   df.triad_country_freq_M$theme_freq_prop_M, na.rm=T))
```

```{r echo=FALSE, paged.print=FALSE, fig.cap ="Correlation between English corpus lexical co-occurrence and US responding, and Vietnamese corpus lexical co-occurrence and Vietnam responding."}
plotE + plotV + plotM
  plot_layout(ncol = 3)
```

We first conducted a Pearson correlation test between corresponding response data and corpus data. Looking at the raw lexical co-occurrence data, proportion of US thematic responding is moderately correlated to English corpus thematic proportion, `r corr.E$full_result`. Proportion of Vietnam thematic responding is moderately correlated to Vietnamese corpus thematic proportion, `r corr.V$full_result`. Proportion of China thematic responding is moderately correlated to Chinese corpus thematic proportion, `r corr.M$full_result`. 
There was a strong similarity signal between the thematic proportions in the English and Vietnamese corpora, `r corr.EV$full_result`, and the thematic proportions in the English and Mandarin corpora, `r corr.EM$full_result`. There was a moderate similarity signal between the proportions in the Mandarin and Vietnamese corpora, `r corr.MV$full_result.`

```{r Pearson correlation between corpus cosine distance and responding, warning=FALSE, include=FALSE}
df.triad_country_cos <- df %>%
  group_by(triad, country, theme_cosine_prop_E, theme_cosine_prop_V, theme_cosine_prop_M) %>%
  summarize(emp_theme_prop = mean(responses_theme))

df.triad_country_cos_E <- df.triad_country_cos %>%
  filter(country == "US")

plotE = ggplot(data = df.triad_country_cos_E, 
        mapping = aes(x = emp_theme_prop, 
                     y = theme_cosine_prop_E,
                     label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)", 
        y = "Proportion Thematic Predicted (English)") + 
   scale_y_reverse() + 
   xlim(0, 1) +
   ylim(0.8, 0.2) + 
  theme(axis.title = element_text(size=rel(0.75)))

df.triad_country_cos_V <- df.triad_country_cos %>%
  filter(country == "Vietnam")

plotV = ggplot(data = df.triad_country_cos_V, 
        mapping = aes(x = emp_theme_prop, 
                     y = theme_cosine_prop_V, 
                     label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (Vietnam)", 
        y = "Proportion Thematic Predicted (Vietnamese)") + 
   scale_y_reverse() +
   xlim(0, 1) +
   ylim(0.8, 0.2) + 
  theme(axis.title = element_text(size=rel(0.75)))

df.triad_country_cos_M <- df.triad_country_cos %>%
  filter(country == "China")

plotM = ggplot(data = df.triad_country_cos_M, 
        mapping = aes(x = emp_theme_prop, 
                     y = theme_cosine_prop_M, 
                     label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (China)", 
        y = "Proportion Thematic Predicted (Mandarin)") + 
   xlim(0, 1) +
   ylim(0.8, 0.2) + 
  theme(axis.title = element_text(size=rel(0.75)))

corr.E_cos = apa_print(cor.test(df.triad_country_cos_E$emp_theme_prop,
                      df.triad_country_cos_E$theme_cosine_prop_E, na.rm=T))

corr.V_cos = apa_print(cor.test(df.triad_country_cos_V$emp_theme_prop,
                      df.triad_country_cos_V$theme_cosine_prop_V, na.rm=T))

corr.M_cos = apa_print(cor.test(df.triad_country_cos_M$emp_theme_prop,
                      df.triad_country_cos_M$theme_cosine_prop_M, na.rm=T))

#correlation between corpora
corr.EV_cos = apa_print(cor.test(df.triad_country_cos_E$theme_cosine_prop_E,
                       df.triad_country_cos_V$theme_cosine_prop_V, na.rm=T))

corr.EM_cos = apa_print(cor.test(df.triad_country_cos_E$theme_cosine_prop_E,
                      df.triad_country_cos_M$theme_cosine_prop_M, na.rm=T))

corr.VM_cos = apa_print(cor.test(df.triad_country_cos_V$theme_cosine_prop_V,
                      df.triad_country_cos_M$theme_cosine_prop_M, na.rm=T))
```

```{r warning = FALSE, echo=FALSE, fig.width = 8, fig.cap="Correlation between English corpus cosine distance and US responding; Vietnamese corpus cosine distance and Vietnam responding; and Mandarin corpus cosine distance and China responding."}
#maybe we should include a note that explains why they are correlate in the opposite direction? 
#either changing the label of y-axis, or use y = 1/theme_cosine_prop_V? 
plotE + plotV + plotM +
  plot_layout(ncol = 3)
```

```{r US-Mandarin, China-English cross correlation, NOT SHOWN, eval=FALSE, warning=FALSE, include=FALSE}
plotUSZH = ggplot(data = df.triad_country_cos_E, 
        mapping = aes(x = emp_theme_prop, 
                     y = theme_cosine_prop_M,
                     label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)", 
        y = "Proportion Thematic Predicted (Mandarin)") + 
   scale_y_reverse() + 
   xlim(0, 1) +
   ylim(0.8, 0.3)

plotCNEN = ggplot(data = df.triad_country_cos_M, 
        mapping = aes(x = emp_theme_prop, 
                     y = theme_cosine_prop_E,
                     label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (China)", 
        y = "Proportion Thematic Predicted (English)") + 
   scale_y_reverse() + 
   xlim(0, 1) +
   ylim(0.8, 0.3)

plotUSZH + plotCNEN +
  plot_layout(ncol = 2)
```

Looking at the cosine distance data using fastText word vectors, proportion of US thematic responding is moderately correlated to English thematic cosine distance proportion, `r corr.E_cos$full_result`. There is a small correlation between proportion of Vietnam thematic responding and the Vietnamese thematic cosine distance proportion, `r corr.V_cos$full_result`. Proportion of China thematic responding is moderately correlated to Mandarin thematic match cosine distance proportion, `r corr.M_cos$full_result`. There were moderate signals of similarity between English and Mandarin (`r corr.EM_cos$full_result`). There was no significant correlation between English and Vietnamese (`r corr.EV_cos$full_result`), and between Vietnamese and Mandarin (`r corr.VM_cos$full_result`).

To test whether variation in lexical co-occurrence can explain differences in similarity judgments between US and Vietnam participants, we compare logistic mixed-effects regression models fit to the thematic responding data from each country separately. We first ask how well each corpus model (English, Vietnamese, or Mandarin) predicts similarity judgments by speakers of the corresponding language (US, Vietnam, or China). To do this, we use a mixed-effects logistic regression to predict triad responses (0=taxonomic or 1=thematic) with corpus prediction (proportion thematic co-occurrence or cosine distance) as a fixed effect and participant and triad as random effects.

#### Raw lexical co-occurrences

```{r US responding predicted by each corpus model raw co-occurrences, warning=FALSE, include=F}
# English 
fit.EN_US = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_US)
fit.EN_US.anova = Anova(fit.EN_US, type = 3)

#Vietnamese
fit.VI_US = glmer(responses_theme ~ theme_freq_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.VI_US)
fit.VI_US.anova = Anova(fit.VI_US, type = 3)

#Mandarin
fit.ZH_US = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_US)
fit.ZH_US.anova = Anova(fit.ZH_US, type = 3)
```

```{r VN responding predicted by each corpus model raw co-occurrences, warning=FALSE, include=F}
# English 
fit.EN_VN = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_VN)
fit.EN_VN.anova = Anova(fit.EN_VN, type = 3)

@@

```{r CN responding predicted by each corpus model raw co-occurrences, warning=FALSE, include=F}
#English
fit.EN_CN = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_CN)
fit.EN_CN.anova = Anova(fit.EN_CN, type = 3)

#Vietnamese
fit.VI_CN = glmer(responses_theme ~ theme_freq_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.VI_CN)
fit.VI_CN.anova = Anova(fit.VI_CN, type = 3)

#Mandarin
fit.ZH_CN = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_CN)
fit.ZH_CN.anova = Anova(fit.ZH_CN, type = 3)
```

TODO: ADD CHINA RESULTS
Both the English (EN) and Vietnamese (VI) corpus are significant predictors of US responding (EN-US model: $\beta$ = `r round(fixef(fit.EN_US)["theme_freq_prop_E"], 2)`, $\chi^2$(`r fit.EN_US.anova["theme_freq_prop_E", "Df"]`) = `r round(fit.EN_US.anova["theme_freq_prop_E", "Chisq"], 2)`, p < .001; VI-US model: $\beta$ = `r round(fixef(fit.VI_US)["theme_freq_prop_V"], 2)`, $\chi^2$(`r fit.VI_US.anova["theme_freq_prop_V", "Df"]`) = `r round(fit.VI_US.anova["theme_freq_prop_V", "Chisq"], 2)`, p < .001).

Similarly, both the English (EN) and Vietnamese (VI) corpus are significant predictors of Vietnam (VN) responding (EN-VN model: $\beta$ = `r round(fixef(fit.EN_VN)["theme_freq_prop_E"], 2)`, $\chi^2$(`r fit.EN_VN.anova["theme_freq_prop_E", "Df"]`) = `r round(fit.EN_VN.anova["theme_freq_prop_E", "Chisq"], 2)`, p < .001; VI-VN model: $\beta$ = `r round(fixef(fit.VI_VN)["theme_freq_prop_V"], 2)`, $\chi^2$(`r fit.VI_VN.anova["theme_freq_prop_V", "Df"]`) = `r round(fit.VI_VN.anova["theme_freq_prop_V", "Chisq"], 2)`, p < .001).

This suggests a strong shared similarity signal across the languages, which is also reflected in the strong correlation between the two corpus models reported above (`r corr.EV$full_result`).

#### Cosine distances of fastText word vectors
```{r US responding predicted by each corpus model cosine differences, warning=FALSE, include=F}
#English
fit.EN_US_cos = glmer(responses_theme ~ theme_cosine_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US"), 
          family="binomial")
summary(fit.EN_US_cos)
fit.EN_US_cos.anova = Anova(fit.EN_US_cos, type = 3)
fit.EN_US_cos.anova

#Vietnamese
fit.VI_US_cos = glmer(responses_theme ~ theme_cosine_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US"), 
          family="binomial")
summary(fit.VI_US_cos)
fit.VI_US_cos.anova = Anova(fit.VI_US_cos, type = 3)
fit.VI_US_cos.anova

#Mandarin
fit.ZH_US_cos = glmer(responses_theme ~ theme_cosine_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US"), 
          family="binomial")
summary(fit.ZH_US_cos)
fit.ZH_US_cos.anova = Anova(fit.ZH_US_cos, type = 3)
fit.ZH_US_cos.anova
```

```{r VN responding predicted by each corpus model cosine differences, warning=FALSE, include=F}
#English
fit.EN_VN_cos = glmer(responses_theme ~ theme_cosine_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam"), 
          family="binomial")
summary(fit.EN_VN_cos)
fit.EN_VN_cos.anova = Anova(fit.EN_VN_cos, type = 3)
fit.EN_VN_cos.anova

#Vietnamese
fit.VI_VN_cos = glmer(responses_theme ~ theme_cosine_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam"), 
          family="binomial")
summary(fit.VI_VN_cos)
fit.VI_VN_cos.anova = Anova(fit.VI_VN_cos, type = 3)
fit.VI_VN_cos.anova

#Mandarin
fit.ZH_VN_cos = glmer(responses_theme ~ theme_cosine_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam"), 
          family="binomial")
summary(fit.ZH_VN_cos)
fit.ZH_VN_cos.anova = Anova(fit.ZH_VN_cos, type = 3)
fit.ZH_VN_cos.anova
```

```{r CN responding predicted by each corpus model cosine differences, warning=FALSE, include=F}
## English
fit.EN_CN_cos = glmer(responses_theme ~ theme_cosine_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China"), 
          family="binomial")
summary(fit.EN_CN_cos)
fit.EN_CN_cos.anova = Anova(fit.EN_CN_cos, type = 3)
fit.EN_CN_cos.anova

#Vietnamese
fit.VI_CN_cos = glmer(responses_theme ~ theme_cosine_prop_V + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China"), 
          family="binomial")
summary(fit.VI_CN_cos)
fit.VI_CN_cos.anova = Anova(fit.VI_CN_cos, type = 3)
fit.VI_CN_cos.anova

#Mandarin
fit.ZH_CN_cos = glmer(responses_theme ~ theme_cosine_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China"), 
          family="binomial")
summary(fit.ZH_CN_cos)
fit.ZH_CN_cos.anova = Anova(fit.ZH_CN_cos, type = 3)
fit.ZH_CN_cos.anova
```

Similar to the results from the analysis with raw lexical co-occurrences, we found that all corpora are significant predictors of all cultural context responding.

>>>>

```{r combine corpus and behavior data, include = F}
source("../2_combine/2_combine.R") # include triad_omits. TODO: Find a way to also save this.
```

```{r incorporate base freq, eval=F, include = F}
df <- read.csv("../../data/data_USCNVN_ENZHVI.csv")
df.basefreq <- read.csv("../../data/corpus/basefreq_EMV.csv")

@@

df <- left_join(df, df.basefreq, by = "triad")

#replacing any 0 raw counts with epsilon
df <- df %>%
  mutate(cue_basefreq_E = ifelse(cue_basefreq_E == 0, 
                                  .Machine$double.eps, cue_basefreq_E),
         tax_basefreq_E = ifelse(tax_basefreq_E == 0, 
                                  .Machine$double.eps, tax_basefreq_E),
         theme_basefreq_E = ifelse(theme_basefreq_E == 0, 
                                  .Machine$double.eps, theme_basefreq_E),
         cue_basefreq_M = ifelse(cue_basefreq_M == 0, 
                                  .Machine$double.eps, cue_basefreq_M), 
         tax_basefreq_M = ifelse(tax_basefreq_M == 0, 
                                  .Machine$double.eps, tax_basefreq_M),
         theme_basefreq_M = ifelse(theme_basefreq_M == 0, 
                                  .Machine$double.eps, theme_basefreq_M),
         cue_basefreq_V = ifelse(cue_basefreq_V == 0, 
                                  .Machine$double.eps, cue_basefreq_V),
         tax_basefreq_V = ifelse(tax_basefreq_V == 0, 
                                  .Machine$double.eps, tax_basefreq_V), 
         theme_basefreq_V = ifelse(theme_basefreq_V == 0, 
                                  .Machine$double.eps, theme_basefreq_V)
         )
```

```{r calculate cue given match and match given cue proportions, include = F, eval = F}
# for now, calculate specifically in terms of which one they have answered? (theme vs match)
# but maybe do for all languages? 
calculate_cue_given_match <- function(responses_theme, language,
                                      tax_basefreq_E, theme_basefreq_E,
                                      theme_frequency_E, tax_frequency_E,
                                      tax_basefreq_M, theme_basefreq_M,
                                      theme_frequency_M, tax_frequency_M,
                                      tax_basefreq_V, theme_basefreq_V,
                                      theme_frequency_V, tax_frequency_V) {
    ifelse(language == "English",
           ifelse(responses_theme == 1, theme_frequency_E / theme_basefreq_E, tax_frequency_E / tax_basefreq_E),
           ifelse(language == "Mandarin",
                  ifelse(responses_theme == 1, theme_frequency_M / theme_basefreq_M, tax_frequency_M / tax_basefreq_M),
                  NA))
}

calculate_match_given_cue <- function(responses_theme, language,
                                      cue_basefreq_E, theme_frequency_E, tax_frequency_E,
                                      cue_basefreq_M, theme_frequency_M, tax_frequency_M,
                                      cue_basefreq_V, theme_frequency_V, tax_frequency_V) {
    ifelse(language == "English",
           ifelse(responses_theme == 1, theme_frequency_E / cue_basefreq_E, tax_frequency_E / cue_basefreq_E),
           ifelse(language == "Mandarin",
                  ifelse(responses_theme == 1, theme_frequency_M / cue_basefreq_M, tax_frequency_M / cue_basefreq_M),
                  NA))
}

df <- df %>% mutate(cue_given_match = calculate_cue_given_match(
                                          responses_theme, language,
                                          tax_basefreq_E, theme_basefreq_E,
                                          theme_frequency_E, tax_frequency_E,
                                          tax_basefreq_M, theme_basefreq_M,
                                          theme_frequency_M, tax_frequency_M,
                                          tax_basefreq_V, theme_basefreq_V,
                                          theme_frequency_V, tax_frequency_V),
                    match_given_cue = calculate_match_given_cue(
                                          responses_theme, language,
                                          cue_basefreq_E, theme_frequency_E, tax_frequency_E,
                                          cue_basefreq_M, theme_frequency_M, tax_frequency_M,
                                          cue_basefreq_V, theme_frequency_V, tax_frequency_V))

write.csv(df, "../../data/data_USCNVN_ENZHVI_direction.csv")
```

## Does match given cue (intuitive) predict behavior better than cue given match (non-intuitive)?

We are trying to figure out if people's process is more well-modeled by match given cue or cue given match. In other words, are people looking at the cue ("cow") and then judging the matches ("chicken" or "grass")? Or are people looking at the matches first ("chicken") and then seeing if it matches with the cue ("cow") well enough?
```{r loading in the csv with direction so we dont need to go through the previous section, include =F}
df <- read.csv("../../data/data_USCNVN_ENZHVI_direction.csv")
```

We need to run the baseline model again (corresponding corpora predicting responding behavior). And then compare that to the new updated models.

### English
Baseline model:
```{r US responding predicted by each corpus model raw co-occurrences, warning=FALSE}
# English 
fit.EN_US = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_US)
fit.EN_US.anova = Anova(fit.EN_US, type = 3)
```

Intuitive model (cue first): 
```{r}
fit.EN_US_cue_first = glmer(responses_theme ~ theme_freq_prop_E + match_given_cue + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_US_cue_first)
```

Comparing cue-first (intuitive) and base model: Cue-first model fits data better than base model.
```{r}
fit.EN_US_dir_compare_cue_first = anova(fit.EN_US, fit.EN_US_cue_first, type = 3)
fit.EN_US_dir_compare_cue_first
```

Non-intuitive model (match first):
```{r}
fit.EN_US_match_first = glmer(responses_theme ~ theme_freq_prop_E + cue_given_match + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "US", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_US_match_first)
```

Comparing match-first (non-intuitive) and base model: Match-first model fits data better than base model.
```{r}
fit.EN_US_dir_compare_match_first = anova(fit.EN_US, fit.EN_US_match_first, type = 3)
fit.EN_US_dir_compare_match_first
```

Comparing match-first (non-intuitive) and cue-first (intuitive) model: Null result?
```{r}
fit.EN_US_dir_compare_cue_vs_match_first = anova(fit.EN_US_cue_first, fit.EN_US_match_first, type = 3)
fit.EN_US_dir_compare_cue_vs_match_first
```

Adding cue given match (cue+match / match) OR match given cue (cue+match / cue) both improve the model significantly over only having thematic frequency prop, even though the additional term does not wipe out the effect of thematic proportion. Comparing between the cue given match and match given cue model yields null result.

```{r VN responding predicted by each corpus model raw co-occurrences, warning=FALSE, include=F, eval=F}
# English 
fit.EN_VN = glmer(responses_theme ~ theme_freq_prop_E + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "Vietnam", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.EN_VN)
fit.EN_VN.anova = Anova(fit.EN_VN, type = 3)

@@

### Mandarin
Baseline model:
```{r CN responding predicted by each corpus model raw co-occurrences, warning=FALSE}
#Mandarin
fit.ZH_CN = glmer(responses_theme ~ theme_freq_prop_M + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_CN)
fit.ZH_CN.anova = Anova(fit.ZH_CN, type = 3)
```

Intuitive model (cue first):
```{r}
fit.ZH_CN_cue_first = glmer(responses_theme ~ theme_freq_prop_M + match_given_cue + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_CN_cue_first)
```

Comparing cue-first (intuitive) and base model: Cue-first model fits data better than base model.
```{r}
fit.ZH_CN_dir_compare_cue_first = anova(fit.ZH_CN, fit.ZH_CN_cue_first, type = 3)
fit.ZH_CN_dir_compare_cue_first
```

Non-intuitive model (match first):
```{r}
fit.ZH_CN_match_first = glmer(responses_theme ~ theme_freq_prop_M + cue_given_match + (1 | subject) + (1 | triad), 
          data = df %>% filter(country == "China", !(triad %in% triads_omit)), 
          family="binomial")
summary(fit.ZH_CN_match_first)
```

Comparing match-first (non-intuitive) and base model: Match-first model fits data better than base model.
```{r}
fit.ZH_CN_dir_compare_match_first = anova(fit.ZH_CN, fit.ZH_CN_match_first, type = 3)
fit.ZH_CN_dir_compare_match_first
```

Comparing match-first (non-intuitive) and cue-first (intuitive) model: Null result?
```{r}
fit.ZH_CN_dir_compare_cue_vs_match_first = anova(fit.ZH_CN_cue_first, fit.ZH_CN_match_first, type = 3)
fit.ZH_CN_dir_compare_cue_vs_match_first

```
Adding cue given match (cue+match / match) OR match given cue (cue+match / cue) both improve the model significantly over only having thematic frequency prop, even though the additional term does not wipe out the effect of thematic proportion. Comparing between the cue given match and match given cue model yields null result.
